---
title: "Assignment 4 - Applying meta-analytic priors"
author: "Astrid Rybner, Kata Molnar, Sofie Rødkjær and Nicole Dwenger"
output:   
  md_document:
    variant: markdown_github
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse, pacman)
pacman::p_load(rethinking, brms, ggplot2, metafor)
pacman::p_load(patchwork)
```

## Assignment 4

In this assignment we do the following:
- we run a Bayesian meta-analysis of pitch variability in ASD, based on previously published literature
- we analyze pitch variability in ASD in two new studies using both a conservative and a meta-analytic prior
- we assess the difference in model quality and estimates using the two priors.

The questions you need to answer are: What are the consequences of using a meta-analytic prior? Evaluate the models with conservative and meta-analytic priors. Discuss the effects on estimates. Discuss the effects on model quality. Discuss the role that meta-analytic priors should have in scientific practice. Should we systematically use them? Do they have drawbacks? Should we use them to complement more conservative approaches? How does the use of meta-analytic priors you suggest reflect the skeptical and cumulative nature of science?

### Step by step suggestions

Step 1: Perform a meta-analysis of pitch variability from previous studies of voice in ASD
- the data is available as Ass4_MetaAnalysisData.tsv
- You should calculate Effect size (cohen's d) and Standard Error (uncertainty in the Cohen's d) per each study, using escalc() from the metafor package (also check the livecoding intro)
- N.B. we're only interested in getting a meta-analytic effect size for the meta-analytic prior (and not e.g. all the stuff on publication bias). See a brms tutorial here: https://vuorre.netlify.com/post/2016/09/29/meta-analysis-is-a-special-case-of-bayesian-multilevel-modeling/ The formula is EffectSize | se(StandardError) ~ 1 + (1 | Paper). Don't forget prior definition, model checking, etc.
- Write down the results of the meta-analysis in terms of a prior for step 2.

```{r}
#calculate cohen's d + se 
#only want meta-analytic prior, so no worrying about other things
#effect size with uncertainty (se) and estimating metaanalaytic mean, given the uncertainty they have, and each paper different 

set.seed(5)

#LOAD DATA
MA_d <- read_delim("Ass4_MetaAnalysisData.tsv", delim = "\t")
#some papers are repeated becasue 2 different studies/populations
#diffferent studies might also work through the same population/data
#dont differentiate between hz and logscale

#CLEAN DATA
#turn pitch var from character into numeric 
MA_d <- MA_d %>% 
  mutate(
    PitchVariabilityASD_Mean = as.numeric(PitchVariabilityASD_Mean),
    PitchVariabilityTD_Mean = as.numeric(PitchVariabilityTD_Mean),
    PitchVariabilityASD_SD = as.numeric(PitchVariabilityASD_SD),
    PitchVariabilityTD_SD = as.numeric(PitchVariabilityTD_SD)
  )

#delete rows with NAs
MA_d <- MA_d %>% subset(!is.na(Paper))

#CALCULATE EFFECT SIZE
#do escalc to get efect size 
#calculate effect size for each study (slab)
#smd = standardised mean difference (cohens d)
#TD-ASD - negative value = ASD bigger pitch variability
MA_d <- escalc(measure = "SMD",
            n1i = TD_N,
            n2i = ASD_N,
            m1i = PitchVariabilityTD_Mean,
            m2i = PitchVariabilityASD_Mean,
            sd1i = PitchVariabilityTD_SD,
            sd2i = PitchVariabilityASD_SD,
            data = MA_d,
            slab = Paper)
            
#calculate standard error sqrt(sd) and renaming effect size
MA_d <- MA_d %>% 
  mutate(StandardError = sqrt(vi)) %>% 
  rename(EffectSize = yi)

#check up: get mean effect size and standard error
summary(MA_d$EffectSize)
summary(MA_d$StandardError)

#CALCULATE META-ANALYTIC EFFECT SIZE
#fixed effect ~ 1 becasue we want a mean  
#varying effects: population varying effect, different data sets
#outcome is not to be taken at face value but take into account the standard error
MA_f <- bf(EffectSize | se(StandardError) ~ 1 + (1|Population))

#get prior
get_prior(MA_f, data = MA_d, family = gaussian)

sd(MA_d$EffectSize, na.rm = T)

#define prior
MA_prior <- c(
  prior(normal(0,1), class = Intercept),
  prior(normal(0,.3), class = sd)
)

#prior model 
MA_m0 <- brm(
  MA_f,
  data = MA_d,
  family = gaussian(),
  prior = MA_prior,
  sample_prior = "only",
  chain = 2,
  cores = 2
)

#prior check
pp_check(MA_m0, nsamples = 100)

#real model
MA_m1 <- brm(
  MA_f,
  data = MA_d,
  family = gaussian(),
  prior = MA_prior,
  sample_prior = T,
  chain = 2,
  cores = 2
)

#posterior check
pp_check(MA_m1, nsamples = 100)

#model summary
summary(MA_m1)

#save values
#MA mean = 0.44
MA_mean = fixef(MA_m1)[1]
#MA sd = 0.1
MA_sd = fixef(MA_m1)[2]
#MA heterogeneity, i.e. difference = 0.32
MA_hetero = 0.32

```

Step 2: Analyse pitch variability in ASD in two new studies for which you have access to all the trials (not just study level estimates)
- the data is available as Ass4_data.csv. Notice there are 2 studies (language us, and language dk), multiple trials per participant, and a few different ways to measure pitch variability (if in doubt, focus on pitch IQR, interquartile range of the log of fundamental frequency)
- Also, let's standardize the data, so that they are compatible with our meta-analytic prior (Cohen's d is measured in SDs).
- Is there any structure in the dataset that we should account for with random/varying effects? How would you implement that? Or, if you don't know how to do bayesian random/varying effects or don't want to bother, is there anything we would need to simplify in the dataset?

```{r}
#how should we analyse them
#how should we include the metaanalytic prior
#what difference does it make
#more robust is pitch variability, but can use others? 
#cohens d measured in sd - so standardise the data
#make sure you look at structure in the data (2studies, trials pp, )

#LOAD DATA
d <- read_csv("Ass4_data.csv", col_types = cols(ID = col_character()))

#SCALE
d <- d %>% 
  mutate(
    PitchVariability = scale(Pitch_IQR)
  )

#CHECK
hist(d$Pitch_IQR)
hist(d$PitchVariability)

#structure in the data 
#different languages (d, e) - Fixed Effect
#id, each participant included several times, becasue going through different trials - Varying effect

```


Step 3: Build a regression model predicting Pitch variability from Diagnosis.
- how is the outcome distributed? (likelihood function). NB. given we are standardizing, and the meta-analysis is on that scale, gaussian is not a bad assumption. Lognormal would require us to convert the prior to that scale.
- how are the parameters of the likelihood distribution distributed? Which predictors should they be conditioned on? Start simple, with Diagnosis only. Add other predictors only if you have the time and energy!
- use a skeptical/conservative prior for the effects of diagnosis. Remember you'll need to motivate it.
- Evaluate model quality. Describe and plot the estimates. 

```{r}
#actually want to build the model predicting pitchvariability from diagnosis 
#whether they is a difference
#describe how the outcome is distirbuted
#describe how the parameters of the likelihood is distributed, in which way is the mean effected by predictiors
#motivate it = need to explain why you are choosing one mean or the other and what consequences that has - before looking at the data
#evaluate model!

#MODELS
#without language: pitch variability is different in the ASD and TD and different individuals will show different variability.
#i.e. hypotheiss that difference in languge makes no difference
NS_f0 <- bf(PitchVariability ~ 1 + Diagnosis + (1|ID))
#with language: 
NS_f1 <- bf(PitchVariability ~ 0 + Language + Language:Diagnosis + (1|ID))
#the two languages have both their own intercept when diagnosis is ASD, and diagnosis is going to make a difference, effect of language by diagnosis

#WITHOUT LANGUAGE
get_prior(NS_f0, d, family = gaussian)
#beta (effect of diganosis TD compared to ASD),
#intercept (average for ASD)
#sd (variability for participant)
#sigma (error we expect that the model makes on a given data point)

NS_prior0 <- c(
  prior(normal(0, .3), class = Intercept), #ASD, average pitch var for asds 
  prior(normal(0, .1), class = b), #expacting a small difference, sceptical prior
  prior(normal(0, .1), class = sd), #no big difference for participants from the mean, not bigger than difference between diagnosis
  prior(normal(.5, .3), class = sigma) #lower boundry of 0, don't expect the error to be 0, is already default 
)

NS_0_priorcheck <- brm(
  NS_f0, 
  d,
  family = gaussian(),
  prior = NS_prior0,
  sample_prior = "only",
  chains = 2,
  cores = 2
) 

pp_check(NS_0_priorcheck, nsamples = 100)

#actual model
NS_m0 <- brm(
  NS_f0, 
  d,
  family = gaussian(),
  prior = NS_prior0,
  sample_prior = T,
  chains = 2,
  cores = 2
) 

pp_check(NS_m0, nsamples = 100)

#output + checks
summary(NS_m0)
#credible evidence that the difference is there, but not that strong and difference is small 
#intercept (asd) at 0.26 and td's are slightly lower (-0.08)

#hyothesis that td has lower pitch var (based on meta analysis)
hypothesis(NS_m0, "DiagnosisTD < 0")
plot(hypothesis(NS_m0, "DiagnosisTD < 0"))
#model has learned a bit

#WITH LANGUAGE
#NS_f1 <- bf(PitchVariability ~ 0 + Language + Language:Diagnosis + (1|ID))
get_prior(NS_f1, d, family = gaussian)
#two intercept by language
#two slopes by language = difference that diagnosis makes in the first and second language

NS_prior1 <- c(
  prior(normal(0, .3), class = b, coef = "Languagedk"), #intercept dk (asd), scaled so close to 0, 
  prior(normal(0, .3), class = b, coef = "Languageus"), #intecept us (asd)
  prior(normal(0, .1), class = b, coef = Languagedk:DiagnosisTD), #effect of diagnosis in dk
  prior(normal(0, .1), class = b, coef = Languageus:DiagnosisTD), #effect of diagnosis in us
  prior(normal(0, .1), class = sd), 
  prior(normal(.5, .3), class = sigma)
)

NS_1_priorcheck <- brm(
  NS_f1, 
  d,
  family = gaussian(),
  prior = NS_prior1,
  sample_prior = "only",
  chains = 2,
  cores = 2
)  

pp_check(NS_1_priorcheck, nsamples = 100)

NS_m1 <- brm(
  NS_f1, 
  d,
  family = gaussian(),
  prior = NS_prior1,
  sample_prior = T,
  chains = 2,
  cores = 2
) 

pp_check(NS_m1, nsamples = 100)

#output + checks
summary(NS_m1)
#effects in languages in opposite directions
#languagedk: danish asd 
#language us: us asd 
#different directions 

#for language dk td has lower var
hypothesis(NS_m1, "Languagedk:DiagnosisTD < 0")
plot(hypothesis(NS_m1, "Languagedk:DiagnosisTD < 0"))
#learned something, more moved toward negative

#for language us td has lower var
hypothesis(NS_m1, "Languageus:DiagnosisTD < 0")
plot(hypothesis(NS_m1, "Languageus:DiagnosisTD < 0"))
#learned something, but opposite: moved toward positive

#difference between languages, crible?
hypothesis(NS_m1, "Languagedk:DiagnosisTD < Languageus:DiagnosisTD")
plot(hypothesis(NS_m1, "Languagedk:DiagnosisTD < Languageus:DiagnosisTD"))
#learned something, effect in dk is smaller than us 

#COMPARE THE MODELS 
NS_m0 <- add_criterion(NS_m0, criterion = "loo", reloo = T)
NS_m1 <- add_criterion(NS_m1, criterion = "loo", reloo = T)
loo_model_weights(NS_m0, NS_m1)
#70% percepnt that the model with language is the best, twice as many chances that the second model is better 

```


Step 4: Now re-run the model with the meta-analytic prior
- Evaluate model quality. Describe and plot the estimates. 

```{r}
#use meta-analytic prior (after checking it!!)

#define prior
NS_prior_informed <- c(
  prior(normal(.2, .3), class = b, coef = "Languagedk"), #mean of both groups
  prior(normal(.2, .3), class = b, coef = "Languageus"), #mean of both groups
  prior(normal(-0.44, .1), class = b, coef = Languagedk:DiagnosisTD), #effect of diagnosis in dk
  prior(normal(-0.44, .1), class = b, coef = Languageus:DiagnosisTD), #effect of diagnosis in us
  prior(normal(0, .1), class = sd), 
  prior(normal(.32, .3), class = sigma)
)

#prior model
NS_prior_informed_check <- brm(
  NS_f1, 
  d,
  family = gaussian(),
  prior = NS_prior1,
  sample_prior = "only",
  chains = 2,
  cores = 2
)  

#prior check
pp_check(NS_prior_informed_check, nsamples = 100)

#real model
NS_m_informed <- brm(
  NS_f1, 
  d,
  family = gaussian(),
  prior = NS_prior_informed,
  sample_prior = T,
  chains = 2,
  cores = 2
) 

#posterior check
pp_check(NS_m_informed, nsamples = 100)

#output + checks
summary(NS_m_informed)

#for language dk td has lower var
hypothesis(NS_m_informed, "Languagedk:DiagnosisTD < 0")
plot(hypothesis(NS_m_informed, "Languagedk:DiagnosisTD < 0"))

#for language us td has lower var
hypothesis(NS_m_informed, "Languageus:DiagnosisTD < 0")
plot(hypothesis(NS_m_informed, "Languageus:DiagnosisTD < 0"))

#difference between languages, 
hypothesis(NS_m_informed, "Languagedk:DiagnosisTD < Languageus:DiagnosisTD")
plot(hypothesis(NS_m_informed, "Languagedk:DiagnosisTD < Languageus:DiagnosisTD"))

NS_m_informed <- add_criterion(NS_m_informed, criterion = "loo", reloo = T)

```


Step 5: Compare the models
- Plot priors and posteriors of the diagnosis effect in both models
- Compare posteriors between the two models
- Compare the two models (LOO)
- Discuss how they compare and whether any of them is best.

```{r}
#plot of priors and posteriors in both models 
#compare estimated out of sample error (loo)

loo_model_weights(NS_m1, NS_m_informed)
loo_compare(NS_m1, NS_m_informed)
#informed model has a lot more chances of being the true one than the uninformed one - lot of reasons to why that is

plot(hypothesis(NS_m1, c("Languagedk:DiagnosisTD < 0", "Languageus:DiagnosisTD < 0")))

plot(hypothesis(NS_m_informed, c("Languagedk:DiagnosisTD < 0", "Languageus:DiagnosisTD < 0")))
```


Step 6: Prepare a nice write up of the analysis and answer the questions at the top.

```{r}

```


Optional step 7: how skeptical should a prior be?
- Try different levels of skepticism and compare them using LOO.

Optional step 8: Include other predictors
- Do age, gender and education improve the model?
- Should they be main effects or interactions?
